Clause Size Reduction with \alluip Learning

### 2020-05-15 CR/paper.tex
fb changes to CR

\newcommand{\fs}[1]{\fontsize{#1}{#1}\selectfont}
\newcommand{\fss}[2]{\fontsize{#1}{#2}\selectfont}
\newcommand{\sat}{SAT\xspace}
\newcommand{\code}[1]{\text{#1}}
\newcommand{\assertionTrail}{\trail}
\newcommand{\oneuip}{1-UIP\xspace}
\newcommand{\alluip}{all-UIP\xspace}
\newcommand{\LBD}{\text{LBD}\xspace}
\newcommand{\gap}{\text{gap}}
\newcommand{\tgap}{t_{\mathit{gap}}}
\newcommand{\ntries}{n_{\mathit{tries}}}
\newcommand{\nsuc}{n_{\mathit{succ}}}
\newcommand{\stablealluip}{\textit{stable-alluip}\xspace}
\newcommand{\tryuiplevel}{\textit{try-uip-level}\xspace}
\newcommand{\allUipPure}{\textit{pure-alluip}\xspace}
\newcommand{\allUipMin}{\textit{min-alluip}\xspace}
\newcommand{\allUipAct}{\textit{alluip-active}}
\newcommand{\allUipIn}{\textit{alluip-inclusive}}
\newcommand{\allUipEx}{\textit{alluip-exclusive}}
\newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
\newcommand{\MapleSeven}{\textit{MapleLCMDist}}
\newcommand{\MapleNine}{\textit{MapleLCMDiscChronoBT-DL-v3} }
\newcommand{\MapleEight}{\textit{MapleLCMDiscChronoBT}}
\newcommand{\MapleNineShort}{\textit{MapleCB-DL} }
\newcommand{\expSAT}{\textit{expMaple-CM-GCBumpOnlyLRB} }
\newcommand{\expSATShort}{\textit{expMaple} }
\newcommand{\cadical}{\textit{CaDiCaL}}
\newcommand{\defaultcadical}{\textit{CaDiCaL-default}}
\newcommand{\MapleIUIPPure}{\text{Maple-\allUipPure}}
\newcommand{\MapleIUIMin}{\text{Maple-\allUipMin}}
\newcommand{\MapleEightShort}{\textit{MapleCB}}
\newcommand{\Set}[2]{\{\,#1\mid#2\,\}}
\newcommand{\Array}[2]{[\,#1\mid#2\,]}
\newcommand{\trail}{\ensuremath{\mathcal{T}}}
\newcommand{\trailIdx}[1]{\ensuremath{\iota(#1)}}
\newcommand{\dlevel}[1]{\ensuremath{\mathit{decLvl}(#1)}}
\newcommand{\dlevels}{\ensuremath{\mathit{decLvls}}}
\newcommand{\var}{\text{var}}
\newcommand{\true}{\textsc{true}\xspace}
\newcommand{\false}{\textsc{false}\xspace}
\newcommand{\reason}[1]{\ensuremath{\mathit{reason}(#1)}}
\newcommand{\resolve}{\bowtie}
\renewcommand{\implies}{\rightarrow}
\newcommand{\ctry}{C_{\mathit{try}}}
\newcommand{\config}{\textit{config}}
\newcommand{\unmarked}{\textbf{unmarked}}
\newcommand{\mrk}{\textbf{mark}}
\setlength{\theorempreskipamount}{3pt}
\setlength{\theorempostskipamount}{3pt}

@@

\begin{document}
\maketitle              % typeset the header of the contribution
% 
\begin{abstract}
    Almost all CDCL \sat solvers use the \oneuip 
    clause learning scheme for learning new clauses from conflicts, 
    and our current understanding
    of \sat solving provides good reasons for using that scheme. In
    particular, the \oneuip scheme yields asserting clauses, and these
    asserting clauses have minimum \LBD among all possible asserting
    clauses. As a result of these advantages, other clause learning
    schemes, like $i$-UIP and \alluip, that were proposed in early
    work are not used in modern solvers. In this paper, we propose a
    new technique for exploiting the \alluip clause learning scheme. 
    Our technique is to employ \alluip learning under the
    constraint that the learnt clause's \LBD does not increase (over the
    minimum established by the \oneuip clause). Our method can learn
    clauses that are significantly smaller than the \oneuip clause while
    preserving the minimum \LBD. Unlike previous clause minimization
    methods, our technique is not limited to learning a sub-clause of
    the \oneuip clause. It can instead learn a completely different
    clause. We show empirically 
    that our method can improve the performance of state of the art solvers.
\end{abstract}

@@

One of the previously proposed schemes was the \alluip scheme
\cite{DBLP:conf/iccad/ZhangMMM01}. In this paper we propose a new way
to exploit the main ideas of this scheme that avoids its main
disadvantage which is that it can learn clauses with higher \LBD
scores. In particular, we propose to use a \alluip like clause
learning scheme to generate smaller learnt clauses which retain the
good properties of standard \oneuip clauses. Our method is related to,
but not the same as, various clause mininimization methods that try to
remove redundant literals from the \oneuip clause yielding a clause that
is a subset of the \oneuip clause, e.g.,
\cite{DBLP:conf/sat/SorenssonB09,DBLP:conf/ijcai/LuoLXML17,DBLP:conf/sat/WieringaH13}.
Our method is orthogonal to clause minimization. In particular, our
approach can learn a clause that is not a subset of the \oneuip clause
but which still serves all of the same purposes as the \oneuip clause.
Various minimization techniques can be applied on top of our method to
remove redundant literals from the clauses we have learnt.

@@

\paragraph{The Trail.}
CDCL \sat solvers maintain a \textbf{trail}, $\trail$, which is a
\textit{non-contradictory, non-redundant sequence of literals} that
have been assigned \true by the solver; i.e.
$\ell\in\trail \implies \lnot\ell \not\in\trail$, and $\trail$
contains no duplicates. Newly assigned literals are added to the end
of the trail, and on backtrack literals are removed from the end of
the trail and unassigned.  If literal $\ell$ is on the trail let
$\trailIdx{\ell}$ denote its index on the trail, i.e,
$\trail[\trailIdx{\ell}] = \ell$. For convenience, we also let
$\trailIdx{\ell} = \trailIdx{\lnot \ell} = \trailIdx{\var(\ell)}$ even
though neither $\lnot \ell$ nor $\var(\ell)$ are actually on
$\trail$. If $x$ and $y$ are both on the trail and
$\trailIdx{x} < \trailIdx{y}$ we say that \textit{$x$ appears before
  $y$ on the trail}.

@@

In most \sat solvers, clause learning is initiated as soon as a clause
is falsified by $\trail$. In this paper we will be concerned with the
subsequent clause learning process which uses $\trail$ to derive a new
clause. We will try to make as few assumptions about how $\trail$ is
managed by the \sat solver as possible. One assumption we will make is
that $\trail$ \emph{remains intact during clause learning} and is only
changed after the new clause is learnt. After the new clause is learnt
$\trail$ will be changed by backtracking.

@@

Ordered trail resolutions resolve unit propagated literals from the
end of the trail to the beginning. Without lost of generality, we can
require that all trail resolutions be ordered.

@@

The relevance of trail resolutions is that all proposed clause
learning schemes we are aware of use trail resolutions to produce
learnt clauses. Furthermore, the commonly used technique for clause
minimization \cite{DBLP:conf/sat/SorenssonB09} is also equivalent to a
trail resolution that yields the minimized clause from the
un-minimized clause. Interestingly, it is standard in \sat solver
implementations to perform resolution going backwards along the
trail. That is, these implementations are typically using ordered
trail resolutions. Observation~\ref{obs:ordered} shows that this in
fact the right way to do this.

@@

\begin{figure}[t!]
{\fss{9pt}{10pt}
\begin{tabular}[t]{|l|c|l|}
  \cline{1-1}\cline{3-3}
  \begin{minipage}[t]{0.48\textwidth}
    \textbf{(a) All Decision Clause\rule{0pt}{1.1\topskip}}
    \begin{algorithmic}[0]
        \State \hspace*{-1em}\textbf{all\_decision}($C_I$)
        \State $C\gets C_I$
        \While{$\{l\,|\, l\in C \land\reason{l} \neq\varnothing\} \neq \emptyset$}
        \State \begin{tabular}{ll}
                 $\ell \gets \mbox{}$ & literal with highest trail index in\\
                                      & $\{l\,|\, l\in C \land\reason{l} \neq\varnothing\}$
               \end{tabular}
        \State $C \gets C\resolve \reason{\lnot \ell}$
        \EndWhile
        \State \textbf{return} $C$
    \end{algorithmic}
  \end{minipage}
  & &
  \begin{minipage}[t]{0.475\textwidth}
      \textbf{(b) Make level $i$ contain a single literal\rule{0pt}{1.1\topskip}}
      \begin{algorithmic}[0]
          \State \hspace*{-1em} \textbf{UIP\_level}($C$, $i$)
          \While{$\left|\left\{\ell \left| \begin{array}{l}\ell \in C\\
                                           \mbox{}\land \reason{l} \neq \varnothing \\
                                           \mbox{}\land \dlevel{l} = i\end{array}\right.\right\}\right| > 1$}
          \State \begin{tabular}{ll}
                   $l \gets \mbox{}$ & literal with highest trail index in\\
                                        & $\left\{\ell \left| \begin{array}{l}\ell \in C
                                          \land \reason{l} \neq \varnothing \\
                                          \mbox{}\land \dlevel{l} = i\end{array}\right.\right\}$
                 \end{tabular}
          \State $C \gets C\resolve \reason{\lnot l}$
          \EndWhile
          \State \textbf{return} $C$\strut
      \end{algorithmic}
  \end{minipage}
  \\\cline{1-1}\cline{3-3}
   \multicolumn{1}{c}{\vspace*{-7pt}}\\\cline{1-1}\cline{3-3}
  \begin{minipage}[t]{0.475\textwidth}
    \begin{minipage}{\textwidth}
        \textbf{(c) First UIP Clause \rule{0pt}{1.1\topskip}}
        \begin{algorithmic}[0]
            \State \hspace*{-1em}\textbf{\oneuip}($C_I$)
            \State $i\gets \max\{\dlevel{l}\, |\, l \in C_I\}$
            \State \textbf{return} \textbf{UIP\_level}($C_I$, i)
        \end{algorithmic}
        \vspace*{5pt}
    \end{minipage}\\[5pt]
    \hspace*{-1.8pt}\rule{1.03\textwidth}{.4pt}\\[5pt]
    \begin{minipage}{\textwidth}
        \textbf{(d) \alluip Clause \rule{0pt}{1.1\topskip}}
        \begin{algorithmic}[0]
          \State \hspace*{-1em}\textbf{\alluip}($C_I$, $i$)
          \State $i = \bigl|\{\dlevel{l} \,|\, l \in \trail\}\bigl|$
          \State \Comment{$i$ is large enough to ensure all levels are UIP}
          \State \textbf{return} \textbf{$i$-UIP}(C, i)

      \end{algorithmic}
      \end{minipage}
  \end{minipage}
 & &
  \begin{minipage}[t]{0.48\textwidth}
      \vspace*{-20pt}
      \textbf{(e) $i$-UIP Clause}
      \begin{algorithmic}[0]
          \State \hspace*{-1em}\textbf{$i$-UIP}($C_I$, $i$)
          \State $C\gets C_I$
          \State $d\gets \max\{\dlevel{l}\,|\, l \in C\}$
          \For{($j \gets 1$; $j\leq i$; $j\gets j+1$)}
          \State \textbf{if} ($d = \varnothing$): \textbf{break}
          \State $C\gets\mbox{}$ \textbf{UIP\_level(C, d)}
          \State $d\gets \max\left\{\dlevel{l} \left|\begin{array}{l}l\in C \\ \mbox{} \land \dlevel{l}<d
                                                  \end{array}\right.\right\}$
          \EndFor
          \State \textbf{return} $C$\strut
      \end{algorithmic}
      \vspace*{3pt}
      \textit{Maximum of an empty set is $\varnothing$ \strut}
  \end{minipage}
\\\cline{1-1}\cline{3-3}
\end{tabular}
}
\caption{Some different clause learning schemes. All use the
      current trail $\trail$ and take as input an initial clause $C_I$
      falsified by $\trail$ at its deepest level.\label{fig:cl_schemes}}
\end{figure}

 %   \floatstyle{boxed}
 %   \restylefloat{figure}

@@

The other schemes must perform more trail resolutions. In fact, all of
these schemes (all-decision, all-UIP, i-UIP) use trail resolutions in
which the \oneuip clause appears. That is, they all must first generate
the \oneuip clause and then continue with further trail resolution
steps. These extra resolution steps can introduce many addition
decision levels into the final clause. Hence, these schemes yield
learnt clauses whose \LBD can be larger, and never smaller, that the
\oneuip clause.

@@

The \alluip scheme will tend to produce small clauses since the
clauses contain at most one literal per decision level. However, the
\alluip clause can have much higher \LBD. Since $\LBD$ is more
important than size, our approach is to use \alluip learning when, and
only when, it succeeds in reducing the size of the clause
\emph{without increasing its \LBD}. The \alluip scheme first computes
the \oneuip clause when it reduces the deepest level to a single UIP
literal. It then proceeds to reduce the shallower levels (see
\alluip's for loop in Figure~\ref{fig:cl_schemes}). So our approach
will start with the \oneuip clause and then try to apply \alluip
learning to reduce other levels to single literals. As noted above,
clause minimization is orthogonal to our approach, so we also first
apply standard clause minimization \cite{DBLP:conf/sat/SorenssonB09}
to the \oneuip clause. That is, our algorithm $\stablealluip$,
Algorithm~\ref{alg:i-uip}, starts with the clause that most \sat
solvers learn from a conflict, a minimized \oneuip clause.

\begin{algorithm}[!t]
\caption{\stablealluip}\label{alg:i-uip}
\begin{algorithmic}[1]
    \Require $C_1$ is minimized \oneuip clause
    \Require \config{} a set of configuration parameters to give different versions \stablealluip.
    \Require $\tgap\geq 0$ is a global parameter, $\ntries$ and $\nsuc$ are used to
         dynamically adjust $\tgap$

@@

\tryuiplevel($C_i$, $i$) attempts to resolve away the literals at
decision level $i$ in the clause $C_i$, i.e., those in the set $L_i$
(line~\ref{ln:litsati1}), in order of decreasing trail index, until
only one literal at level $i$ remains. If the resolution step will not
introduce any new decision levels (line~\ref{ln:nonewdlevel}), it is
performed updating $\ctry$. In addition, all new literals added to
$\ctry$ at level $i$ are added to $L_i$. These are the literals at
level $i$ in the reason clause.

On the other hand, if the resolution step would introduce new decision
levels (line~\ref{ln:newdlevel}) then there are two options.  The
first option we call $\allUipPure$. With $\allUipPure$ we abort our
attempt to UIP this level and return the clause with level $i$
unchanged. In the second option, called $\allUipMin$, we continue
without performing the resolution, keeping the current literal $p$ in
$\ctry$. $\allUipMin$ then continues to try to resolve away the other
literals in $L_i$ (note that $p$ is no longer in $L_i$) until $L_i$ is
reduced to a single literal. Hence, $\allUipMin$ can leave multiple
literals at level $i$---all of those with reasons containing new
levels along with one other.\footnote{Since the sole remaining literal
  $u\in L_i$ is at a lower trail index than all of the other literals
  there is no point in trying to resolve away $u$---either it will be
  the decision literal for level $i$ having no reason, or its reason
  will contain at least one other literal at level $i$.} Observe that
the number of literals at level $i$ can not be increased after
processing it with \allUipPure. \allUipMin can, however, potentially
increase the number of literals at level $i$. In resolving away a
literal $l$ at level $i$, more literals might be introduced into level
$i$, and some of these might not be removable by \allUipMin if their
reasons contain new levels. However, both \allUipPure and \allUipMin
can increase the number of literals at levels less that $i$ as new
literals can be introduced into those levels when the literals at
level $i$ are resolved away. These added literals at the lower levels
might not be removable from the clause, and thus both methods might
yield a longer clause than the inputted \oneuip clause.

@@

\vspace*{-3pt}
\begin{example}
    Consider the trail $\assertionTrail = \ldots$, $\ell_1$, $a_2$, $b_2$,
    $c_2$,$d_2$, $\ldots$, $\ldots$, $e_5$, $f_5$, $g_5$,
    $h_6$, $i_6$, $j_6$, $k_6$, $\ldots$, $m_{10}$,
    $\ldots$ where the subscript indicates the decision level of each
    literal and the literals are in order of increasing trail
    index. 
    \[\begin{array}{|l|l|l|}\hline
      C_{a} = \varnothing 
      & C_{b} =  (b_2, \lnot \ell_3, \lnot a_2)
      & C_{c} =  (c_2, \lnot a_2, \lnot b_2) \\\hline
      C_{d} =  (d_2, \lnot b_2, \lnot c_2) 
      & C_{\ell} =\varnothing
      & C_{e} =\varnothing \\\hline
      C_{f} =  (f_5, \lnot e_5,\lnot \ell_1)
      & C_{g} = (g_5, \lnot a_2, \lnot f_5)
      & C_{h} =\varnothing \\\hline
      C_{i} =  (i_6, \lnot e_5, \lnot h_6)
      & C_{j} = (j_6, \lnot f_5, \lnot i_6 )
      & C_{k} = (k_6, \lnot f_5, \lnot j_6) \\\hline
      \end{array}
  \]
  Let the clauses $C_x$, show above, denote the reason clause for
  literal $x_i$.  Suppose \oneuip learning yields the clause
  $C_1 = (\neg{m_{10}}$, $\neg{k_6}$, $\neg{j_6}$, $\neg{i_6}$,
  $\neg{h_6}$, $\neg{g_5}$, $\neg{d_2}$, $\neg{c_2})$ where
  $\neg{m_{10}}$ is the UIP from the conflicting
  level. $\stablealluip$ first tries to find the UIP for level 6 by
  resolving $C_1$ with $C_k$, $C_j$ and then $C_i$ producing the
  clause $C^* = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{g_5}$, $\neg{f_5}$,
  $\neg{e_5}$, $\neg{d_2}$, $\neg{c_2})$ where $\neg{h_6}$ is the UIP
  for level 6.

>>>>

\newcommand{\fs}[1]{\fontsize{#1}{#1}\selectfont}
\newcommand{\fss}[2]{\fontsize{#1}{#2}\selectfont}
\newcommand{\sat}{SAT\xspace}
\newcommand{\code}[1]{\text{#1}}
\newcommand{\assertionTrail}{\trail}
\newcommand{\oneuip}{1-UIP\xspace}
\newcommand{\alluip}{all-UIP\xspace}
\newcommand{\LBD}{\text{LBD}\xspace}
\newcommand{\gap}{\text{gap}}
\newcommand{\tgap}{t_{\mathit{gap}}}
\newcommand{\ntries}{n_{\mathit{tries}}}
\newcommand{\nsuc}{n_{\mathit{succ}}}
\newcommand{\stablealluip}{\textit{stable-alluip}\xspace}
\newcommand{\tryuiplevel}{\textit{try-uip-level}\xspace}
\newcommand{\allUipPure}{\textit{pure-alluip}\xspace}
\newcommand{\allUipMin}{\textit{min-alluip}\xspace}
\newcommand{\allUipAct}{\textit{alluip-active}}
\newcommand{\allUipIn}{\textit{alluip-inclusive}}
\newcommand{\allUipEx}{\textit{alluip-exclusive}}
\newcommand{\MapleBase}{\textit{MapleCOMSPS-LRB}}
\newcommand{\MapleSeven}{\textit{MapleLCMDist}}
\newcommand{\MapleNine}{\textit{MapleLCMDiscChronoBT-DL-v3} }
\newcommand{\MapleEight}{\textit{MapleLCMDiscChronoBT}}
\newcommand{\MapleNineShort}{\textit{MapleCB-DL} }
\newcommand{\expSAT}{\textit{expMaple-CM-GCBumpOnlyLRB} }
\newcommand{\expSATShort}{\textit{expMaple} }
\newcommand{\cadical}{\textit{CaDiCaL}}
\newcommand{\defaultcadical}{\textit{CaDiCaL-default}}
\newcommand{\MapleIUIPPure}{\text{Maple-\allUipPure}}
\newcommand{\MapleIUIMin}{\text{Maple-\allUipMin}}
\newcommand{\MapleEightShort}{\textit{MapleCB}}
\newcommand{\Set}[2]{\{\,#1\mid#2\,\}}
\newcommand{\Array}[2]{[\,#1\mid#2\,]}
\newcommand{\trail}{\ensuremath{\mathcal{T}}}
\newcommand{\trailIdx}[1]{\ensuremath{\iota(#1)}}
\newcommand{\dlevel}[1]{\ensuremath{\mathit{decLvl}(#1)}}
\newcommand{\dlevels}{\ensuremath{\mathit{decLvls}}}
\newcommand{\var}{\text{var}}
\newcommand{\true}{\textsc{true}\xspace}
\newcommand{\false}{\textsc{false}\xspace}
\newcommand{\reason}[1]{\ensuremath{\mathit{reason}(#1)}}
\newcommand{\resolve}{\bowtie}
\renewcommand{\implies}{\rightarrow}
\newcommand{\ctry}{C_{\mathit{try}}}
\newcommand{\config}{\textit{config}}
\newcommand{\unmarked}{\textbf{unmarked}}
\newcommand{\mrk}{\textbf{mark}}
\setlength{\theorempreskipamount}{3pt}
\setlength{\theorempostskipamount}{3pt}

@@

\begin{document}
\maketitle              % typeset the header of the contribution
% 
\begin{abstract}
    Almost all CDCL \sat solvers use the \oneuip clause learning
    scheme for learning new clauses from conflicts, and our current
    understanding of \sat solving provides good reasons for using that
    scheme. In particular, the \oneuip scheme yields asserting
    clauses, and these asserting clauses have minimum \LBD among all
    possible asserting clauses. As a result of these advantages, other
    clause learning schemes, like $i$-UIP and \alluip, that were
    proposed in early work are not used in modern solvers. In this
    paper, we propose a new technique for exploiting the \alluip
    clause learning scheme.  Our technique is to employ \alluip
    learning under the constraint that the learnt clause's \LBD does
    not increase (over the minimum established by the \oneuip
    clause). Our method can learn clauses that are significantly
    smaller than the \oneuip clause while preserving the minimum
    \LBD. Unlike previous clause minimization methods, our technique
    is not limited to learning a sub-clause of the \oneuip clause. We
    show empirically that our method can improve the performance of
    state of the art solvers.
\end{abstract}

@@

One of the previously proposed schemes was the \alluip scheme
\cite{DBLP:conf/iccad/ZhangMMM01}. In this paper we propose a new way
to exploit the main ideas of this scheme that avoids its main
disadvantage which is that it can learn clauses with higher \LBD
scores. In particular, we propose to use a \alluip like clause
learning scheme to generate smaller learnt clauses which retain the
good properties of standard \oneuip clauses. Our method is related to,
but not the same as, various clause minimization methods that try to
remove redundant literals from the \oneuip clause yielding a clause that
is a subset of the \oneuip clause, e.g.,
\cite{DBLP:conf/sat/SorenssonB09,DBLP:conf/ijcai/LuoLXML17,DBLP:conf/sat/WieringaH13}.
Our method is orthogonal to clause minimization. In particular, our
approach can learn a clause that is not a subset of the \oneuip clause
but which still serves all of the same purposes as the \oneuip clause.
Clause minimization techniques can be applied on top of our method 
to remove redundant literals.

@@

\vspace*{3pt}\noindent\textit{The Trail.}
CDCL \sat solvers maintain a \textbf{trail}, $\trail$, which is a
\textit{non-contradictory, non-redundant sequence of literals} that
have been assigned \true by the solver; i.e.
$\ell\in\trail \implies \lnot\ell \not\in\trail$, and $\trail$
contains no duplicates. Newly assigned literals are added to the end
of the trail, and on backtrack literals are removed from the end of
the trail and unassigned.  If literal $\ell$ is on the trail let
$\trailIdx{\ell}$ denote its index on the trail, i.e,
$\trail[\trailIdx{\ell}] = \ell$. For convenience, we also let
$\trailIdx{\ell} = \trailIdx{\lnot \ell} = \trailIdx{\var(\ell)}$ even
though neither $\lnot \ell$ nor $\var(\ell)$ are actually on
$\trail$. If $x$ and $y$ are both on the trail and
$\trailIdx{x} < \trailIdx{y}$ we say that \textit{$x$ appears before
  $y$ on the trail}.

@@

In most \sat solvers, clause learning is initiated as soon as a clause
is falsified by $\trail$. In this paper we will be concerned with the
subsequent clause learning process which uses $\trail$ to derive a new
clause. We will try to make as few assumptions about how $\trail$ is
managed by the \sat solver as possible. One assumption we will make is
that $\trail$ \emph{remains intact during clause learning} and is only
changed after the new clause is learnt (by backtracking).

@@

Ordered trail resolutions resolve unit propagated literals from the
end of the trail to the beginning. W.l.o.g we can require that all trail
resolutions be ordered.

@@

The relevance of trail resolutions is that all proposed clause
learning schemes we are aware of use trail resolutions to produce
learnt clauses. Furthermore, the commonly used technique for clause
minimization \cite{DBLP:conf/sat/SorenssonB09} is also equivalent to a
trail resolution that yields the minimized clause from the
un-minimized clause. Interestingly, it is standard in \sat solver
implementations to perform resolution going backwards along the
trail. That is, these implementations are typically using ordered
trail resolutions. Observation~\ref{obs:ordered} shows that this is
correct.

@@

{\setlength{\textfloatsep}{-10pt plus 1.0pt minus 2.0pt}
\begin{figure}[t!]
{\fss{8pt}{9pt}
\setlength{\tabcolsep}{0.02em}
\begin{tabular}[t]{|c|c|c|}
  \cline{1-1}\cline{3-3}
  \begin{minipage}[t]{0.5\textwidth}
    \textbf{(a) All Decision Clause\rule{0pt}{1.1\topskip}}
    \begin{algorithmic}[0]
        \State \hspace*{-.7em}\textbf{all-decision}($C_I$)
        \State $C\gets C_I$
        \While{$\{l\,|\, l\in C \land\reason{l} \neq\varnothing\} \neq \emptyset$}
        \State \begin{tabular}{ll}
                 $\ell \gets \mbox{}$ & literal with highest trail index in\\
                                      & $\{l\,|\, l\in C \land\reason{l} \neq\varnothing\}$
               \end{tabular}
        \State $C \gets C\resolve \reason{\lnot \ell}$
        \EndWhile
        \State \textbf{return} $C$
    \end{algorithmic}
  \end{minipage}
  & &
  \begin{minipage}[t]{0.5\textwidth}
      \textbf{(b) Make level $i$ contain a single literal\rule{0pt}{1.1\topskip}}
      \begin{algorithmic}[0]
          \State \hspace*{-.7em} \textbf{UIP-level}($C$, $i$)
          \While{$\left|\left\{\ell \left| \begin{array}{l}\ell \in C\\
                                           \mbox{}\land \reason{l} \neq \varnothing \\
                                           \mbox{}\land \dlevel{l} = i\end{array}\right.\right\}\right| > 1$}
          \State \begin{tabular}{ll}
                   $l \gets \mbox{}$ & literal with highest trail index in\\
                                        & $\left\{\ell \left| \begin{array}{l}\ell \in C
                                          \land \reason{l} \neq \varnothing \\
                                          \mbox{}\land \dlevel{l} = i\end{array}\right.\right\}$
                 \end{tabular}
          \State $C \gets C\resolve \reason{\lnot l}$
          \EndWhile
          \State \textbf{return} $C$\strut
      \end{algorithmic}
  \end{minipage}
  \\\cline{1-1}\cline{3-3}
   \multicolumn{1}{c}{\vspace*{-7pt}}\\\cline{1-1}\cline{3-3}
  \begin{minipage}[t]{0.5\textwidth}
    \begin{minipage}{\textwidth}
        \textbf{(c) First UIP Clause \rule{0pt}{1.1\topskip}}
        \begin{algorithmic}[0]
            \State \hspace*{-.7em}\textbf{\oneuip}($C_I$)
            \State $i\gets \max\{\dlevel{l}\, |\, l \in C_I\}$
            \State \textbf{return} \textbf{UIP-level}($C_I$, i)
        \end{algorithmic}
        \vspace*{5pt}
    \end{minipage}\\[5pt]
    \hspace*{-.6pt}\rule{1.005\textwidth}{.4pt}\\[5pt]
    \begin{minipage}{\textwidth}
        \textbf{(d) \alluip Clause \rule{0pt}{1.1\topskip}}
        \begin{algorithmic}[0]
          \State \hspace*{-.7em}\textbf{\alluip}($C_I$, $i$)
          \State $i = \bigl|\{\dlevel{l} \,|\, l \in \trail\}\bigl|$
          \State \Comment{{\fs{7pt}$i$ is large enough to ensure all levels are UIP}}
          \State \textbf{return} \textbf{$i$-UIP}(C, i)

      \end{algorithmic}
      \end{minipage}
  \end{minipage}
 & &
  \begin{minipage}[t]{0.5\textwidth}
      \vspace*{-20pt}
      \textbf{(e) $i$-UIP Clause}
      \begin{algorithmic}[0]
          \State \hspace*{-.7em}\textbf{$i$-UIP}($C_I$, $i$)
          \State $C\gets C_I$
          \State $d\gets \max\{\dlevel{l}\,|\, l \in C\}$
          \For{($j \gets 1$; $j\leq i$; $j\gets j+1$)}
          \State \textbf{if} ($d = \varnothing$): \textbf{break}
          \State $C\gets\mbox{}$ \textbf{UIP-level(C, d)}
          \State $d\gets \max\left\{\dlevel{l} \left|\begin{array}{l}l\in C \\ \mbox{} \land \dlevel{l}<d
                                                  \end{array}\right.\right\}$
          \EndFor
          \State \textbf{return} $C$\strut
      \end{algorithmic}
      \vspace*{3pt}
      \textit{Maximum of an empty set is $\varnothing$ \strut}
  \end{minipage}
\\\cline{1-1}\cline{3-3}
\end{tabular}
}
\vspace*{-5pt}
\caption{Some different clause learning schemes. All use the
      current trail $\trail$ and take as input an initial clause $C_I$
      falsified by $\trail$ at its deepest level.\label{fig:cl_schemes}}
\end{figure}
}
 %   \floatstyle{boxed}
 %   \restylefloat{figure}

@@

The other schemes must perform more trail resolutions. In fact, all of
these schemes (all-decision, all-UIP, i-UIP) use trail resolutions in
which the \oneuip clause appears. That is, they all must first
generate the \oneuip clause and then continue with further trail
resolution steps. These extra resolution steps can introduce many
addition decision levels into the final clause. Hence, these schemes
learn clauses with \LBD at least as large as the \oneuip clauses.

@@

The \alluip scheme will tend to produce small clauses since the
clauses contain at most one literal per decision level. However, the
\alluip clause can have much higher \LBD. Since $\LBD$ is more
important than size, our approach is to use \alluip learning when, and
only when, it succeeds in reducing the size of the clause
\emph{without increasing its \LBD}. The \alluip scheme first computes
the \oneuip clause when it reduces the deepest level to a single UIP
literal. It then proceeds to reduce the shallower levels (see
\alluip's for loop in Figure~\ref{fig:cl_schemes}). So our approach
will start with the \oneuip clause and then try to apply \alluip
learning to reduce other levels to single literals. As noted above,
clause minimization is orthogonal to our approach, so we also first
apply standard clause minimization \cite{DBLP:conf/sat/SorenssonB09}
to the \oneuip clause. That is, our algorithm $\stablealluip$
(Algorithm~\ref{alg:i-uip}), starts with the clause that most \sat
solvers learn from a conflict, a minimized \oneuip clause.

\begin{algorithm}[!t]
\caption{\stablealluip}\label{alg:i-uip}
\fss{9pt}{10pt}
\begin{algorithmic}[1]
    \Require $C_1$ is minimized \oneuip clause
    \Require \config{} a set of configuration parameters to give different versions \stablealluip.
    \Require $\tgap\geq 0$ is a global parameter, $\ntries$ and $\nsuc$ are used to
         dynamically adjust $\tgap$

@@

\tryuiplevel($C_i$, $i$) attempts to resolve away the literals at
decision level $i$ in the clause $C_i$, i.e., those in the set $L_i$
(line~\ref{ln:litsati1}), in order of decreasing trail index, until
only one literal at level $i$ remains. If the resolution step will not
introduce any new decision levels (line~\ref{ln:nonewdlevel}), it is
performed updating $\ctry$. In addition, all new literals added to
$\ctry$ at level $i$ are added to $L_i$.

On the other hand, if the resolution step would introduce new decision
levels (line~\ref{ln:newdlevel}) then there are two options.  The
first option we call $\allUipPure$. With $\allUipPure$ we abort our
attempt to UIP this level and return the clause with level $i$
unchanged. In the second option, called $\allUipMin$, we continue
without performing the resolution, keeping the current literal $p$ in
$\ctry$. $\allUipMin$ then continues to try to resolve away the other
literals in $L_i$ (note that $p$ is no longer in $L_i$) until $L_i$ is
reduced to a single literal. Hence, $\allUipMin$ can leave multiple
literals at level $i$---all of those with reasons containing new
levels along with one other.\footnote{Since the sole remaining literal
  $u\in L_i$ is at a lower trail index than all of the other literals
  there is no point in trying to resolve away $u$---either it will be
  the decision literal for level $i$ having no reason, or its reason
  will contain at least one other literal at level $i$.} Observe that
the number of literals at level $i$ can not be increased after
processing it with \allUipPure. \allUipMin can, however, potentially
increase the number of literals at level $i$. In resolving away a
literal $l$ at level $i$, more literals might be introduced into level
$i$, and some of these might not be removable by \allUipMin if their
reasons contain new levels. However, both \allUipPure and \allUipMin
can increase the number of literals at levels less that $i$ as new
literals can be introduced into those levels when the literals at
level $i$ are resolved away. These added literals at the lower levels
might not be removable from the clause, and thus both methods 
can yield a longer clause than the input \oneuip clause.

@@

%\vspace*{-3pt}
\begin{example}
    Consider the trail $\assertionTrail = \ldots$, $\ell_1$, $a_2$, $b_2$,
    $c_2$,$d_2$, $\ldots$, $\ldots$, $e_5$, $f_5$, $g_5$,
    $h_6$, $i_6$, $j_6$, $k_6$, $\ldots$, $m_{10}$,
    $\ldots$ where the subscript indicates the decision level of each
    literal and the literals are in order of increasing trail
    index. 
    \[\fss{9pt}{9pt}
        \begin{array}{|l|l|l|}\hline
      C_{a} = \varnothing 
      & C_{b} =  (b_2, \lnot \ell_3, \lnot a_2)
      & C_{c} =  (c_2, \lnot a_2, \lnot b_2) \\\hline
      C_{d} =  (d_2, \lnot b_2, \lnot c_2) 
      & C_{\ell} =\varnothing
      & C_{e} =\varnothing \\\hline
      C_{f} =  (f_5, \lnot e_5,\lnot \ell_1)
      & C_{g} = (g_5, \lnot a_2, \lnot f_5)
      & C_{h} =\varnothing \\\hline
      C_{i} =  (i_6, \lnot e_5, \lnot h_6)
      & C_{j} = (j_6, \lnot f_5, \lnot i_6 )
      & C_{k} = (k_6, \lnot f_5, \lnot j_6) \\\hline
      \end{array}
  \]
  Let the clauses $C_x$, show above, denote the reason clause for
  literal $x_i$.  Suppose \oneuip learning yields the clause
  $C_1 = (\neg{m_{10}}$, $\neg{k_6}$, $\neg{j_6}$, $\neg{i_6}$,
  $\neg{h_6}$, $\neg{g_5}$, $\neg{d_2}$, $\neg{c_2})$ where
  $\neg{m_{10}}$ is the UIP from the conflicting
  level. $\stablealluip$ first tries to find the UIP for level 6 by
  resolving $C_1$ with $C_k$, $C_j$ and then $C_i$ producing the
  clause $C^* = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{g_5}$, $\neg{f_5}$,
  $\neg{e_5}$, $\neg{d_2}$, $\neg{c_2})$ where $\neg{h_6}$ is the UIP
  for level 6.
